---
title: Timings across different  
author: 
- name: Aaron Lun
  email: infinite.monkeys.with.keyboards@gmail.com
output:
  html_document
date: "Revised: September 25, 2020"
package: yaumap
vignette: >
  %\VignetteIndexEntry{Animated UMAPs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

# Timing UMAP runs

## Loading the data

Using the MNIST data:

```{r}
library(snedata)

if (file.exists("mnist_fashion.rds")) {
    mnist <- readRDS("mnist_fashion.rds")
} else {
    mnist <- download_fashion_mnist()
    saveRDS(mnist, file="mnist_fashion.rds")
}

y <- as.matrix(mnist[,-(ncol(mnist)- 0:1)])
storage.mode(y) <- "double"
```

```{r, eval=FALSE}
library(snedata)

if (file.exists("mnist_digits.rds")) {
    mnist <- readRDS("mnist_digits.rds")
} else {
    mnist <- download_mnist()
    saveRDS(mnist, file="mnist_digits.rds")
}

y <- as.matrix(mnist[,-ncol(mnist)])
storage.mode(y) <- "double"
```

## Single thread

We run the standard UMAP algorithm with Annoy-based neighbor detection and using a single thread.

```{r}
library(yaumap)
system.time(out <- umap_from_matrix(y))
plot(out[,1], out[,2], col=mnist$Label)
```

For comparison, we use **uwot** with the same settings.

```{r}
library(uwot)
system.time(ref <- umap(y, n_epochs=500, n_threads=1))
plot(ref[,1], ref[,2], col=mnist$Label)
```

## Multiple threads

Reference **uwot** implementation:

```{r}
system.time(ref <- umap(y, n_epochs=500, n_threads=6))
plot(ref[,1], ref[,2], col=mnist$Label)
```

Only parallelizing various parts of the initialization.

```{r}
system.time(out <- umap_from_matrix(y, nthreads=6))
plot(out[,1], out[,2], col=mnist$Label)
```

Parallelizing everything in batch mode.

```{r}
system.time(out <- umap_from_matrix(y, nthreads=6, batch=TRUE))
plot(out[,1], out[,2], col=mnist$Label)
```

# Session info {-}

```{r}
sessionInfo()
```
